---
title: "Samsung Data Wraggling Project"
author: "Gaurang Desai"
date: "February 6, 2016"
output: html_document
---

```{r}

if (!require(dplyr)) {
    install.packages("dplyr")
    require(dplyr)
}

if (!require(tidyr)) {
    install.packages("tidyr")
    require(tidyr)
}

if (!require(data.table)) {
    install.packages("data.table")
    require(data.table)
}

```
**Data Set Information:**

The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain.

Check the README.txt file for further details about this dataset. 

** Load txt files in dataset **

```{r}

x <- read.table("test/X_test.txt")
y <- read.table("test/y_test.txt")
subject <- read.table("test/subject_test.txt")
feature <- read.table("features.txt")

```

Append data from train

```{r}

x <- rbind(x, read.table("train/X_train.txt"))
y <- rbind(y, read.table("train/y_train.txt"))
subject <- rbind(subject, read.table("train/subject_train.txt"))

```

**Select mean and standard columns : **

First assign column name from features

```{r}
colnames(x) <- c(make.names(feature$V2, unique = TRUE, allow_ = TRUE))
```

Select columns that contains "mean" and "standard"

```{r}
mean_std <- x %>% select(contains("mean"), contains("std"))
```

** Create Activity Table : **

```{r}

ActivityLabel = c(1,2,3,4,5,6)
ActivityName = c("WALKING","WALKING_UPSTAIRS","WALKING_DOWNSTAIRS","SITTING","STANDING","LAYING")
Activity = data.frame(ActivityLabel,ActivityName)

```

Join activity with features

```{r}

xy <- bind_cols(x,y)
setnames(xy,"V1","ActivityLabel")
xy_name <- left_join(xy, Activity)
xy_subject_name <- bind_cols(subject,xy_name)

```

Now use tidy gather function to tidy up the data.

```{r}
xy_subject_name_tidy <- gather(xy_subject_name, "feature", "value", 2:562)
```

Now group by activity, subject and mean value

```{r}

tidy_data <- 
  xy_subject_name_tidy %>% 
  group_by(ActivityName, V1, feature) %>% 
  summarise( mean = mean(value))

```

** Export Data **

```{r}
write.table(xy_subject_name, "subject_activity_features.txt", sep = "\t")
write.table(mean_std, "mean_std.txt", sep = "\t")
write.table(tidy_data, "tidy_data.txt", sep = "\t")
write.table(xy_subject_name_tidy, "tidy_data_by_activity_name.txt", sep = "\t")
write.table(y, "tidy_data.txt", sep = "\t")
write.table(y, "y.txt", sep = "\t")
write.table(x, "x.txt", sep = "\t")


```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
